---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

We start by unziping the file, reading data and having a look at it

```{r reading}
unzip("activity.zip")
read.csv("activity.csv") -> activity.df
str(activity.df)
```


## What is mean total number of steps taken per day?

First we create a data frame, that summarizes the number of steps by days using *dplyr* package

```{r databydays}
library(dplyr)
step_days <- activity.df %>%
    group_by(date) %>%
    summarize(steps = sum(steps, na.rm=TRUE))

step_days
```

Then we plot the data, using *ggplot2*
```{r plotbydays}
library(ggplot2)
qplot(steps, data = step_days, xlab = "Steps per day")
```

Finally, we report the mean and median number of steps per day
```{r m_days}
mean(step_days$steps) -> mean_steps
median(step_days$steps) -> median_steps

```

The mean is `r mean_steps` and median is `r median_steps`

## What is the average daily activity pattern?

We start by creating a data frame with data by intervals

```{r databyintervals}
step_intervals <- activity.df %>%
    group_by(interval) %>%
    summarize(steps = mean(steps, na.rm=TRUE))

step_intervals
```

Now we plot the computed data

```{r plotbyintervals}
qplot(interval, steps, data=step_intervals, geom = "line", ylab="steps on average")
```

Finaly, let's determine the interval with maximum number of steps

```{r m_intervals}
max(step_intervals$steps) -> maxsteps
step_intervals$steps==maxsteps
filter(step_intervals, steps==maxsteps)
```



## Imputing missing values

First let's calculate the number of rows with missing values in the dataset

``` {r NAs}
sum(!complete.cases(activity.df))
```

Now we fill the NAs with the mean steps for that inverval
``` {r NAfilling}
act_noNA <- activity.df
for (i in 1:nrow(act_noNA)) {
    if (is.na(act_noNA$steps[i])) {
        act_noNA$interval[i] -> inter
        step_intervals[step_intervals$interval==inter, 2] -> steps
        act_noNA$steps[i] <- as.double(steps)
    }
}

```

Now let's repeat the previous analyses without the missing data
```{r noNA}
step_days_noNA <- act_noNA %>%
    group_by(date) %>%
    summarize(steps = sum(steps, na.rm=TRUE))

qplot(steps, data = step_days_noNA, xlab = "Steps per day")
```

Finally, we report the mean and median number of steps per day without NA
```{r m_days_noNA}
mean(step_days_noNA$steps) -> mean_steps_noNA
median(step_days_noNA$steps) -> median_steps_noNA

```

The corrected mean is `r mean_steps_noNA` and corrected median is `r median_steps_noNA`
Removing NAs did not change the median, but increased the mean

## Are there differences in activity patterns between weekdays and weekends?
